Parsing Postfix log files

### 2007-10-13 doc/logparser.tex
Improved AWStats review. Improved Logrep review. Improved conclusion of "review of other parsers" section.

XXX THIS NEEDS TO BE EXPANDED\@; EXPLAIN WHY THEY'RE NOT GOOD ENOUGH

@@

    \item [AWStats] \textit{AWStats is a free powerful and featureful tool
        that generates advanced web, streaming, ftp or mail server
        statistics, graphically.\/}

        AWStats will produce simple graphs for many different services, but
        that restricts it to supporting the Lowest Common Denominator.  The
        data it will extract from an MTA log file is limited to: \newline
        \hspace*{2em} time2, email, email\_r, host, host\_r, method, url,
        code and bytesd.  \newline  There does not seem to be an
        explanation of those fields in the documentation.  As AWStats is a
        generic log processor it can safely be assumed that it has
        no special purpose code to deal with the intricacies of Postfix
        logs.

        \url{http://awstats.sourceforge.net/} \newline
        \url{http://awstats.sourceforge.net/awstats.mail.html} \newline
        \url{http://awstats.sourceforge.net/docs/awstats_faq.html#MAIL}
        \newline (Last checked 2007/08/13.)

@@

    \item [Logrep] \textit{Logrep is a secure multi-platform framework for
        the collection, extraction, and presentation of information from
        various log files.\/}

        The parsing Logrep performs is extremely basic, counting the number
        of different processes executed, from address, to addresses, number
        of recipients, bytes transferred and delay.  \newline
        \url{http://www.itefix.no/phpws/index.php} \newline (Last checked
        2007/08/13.)

    \item [Log Mail Analyser] has already been discussed in
        section~\ref{prior art}.

@@

While there are other programs available which perform basic Postfix log
parsing (some to a greater level of intricacy and detail than others), none
attempt to correlate log lines by pid or queueid to produce an overall
record of the journey each mail traverses through Postfix.  Most other
parsers immediately generate a report and discard the data extracted from
the logs; those which don't discard the data retain it in a format
inaccessible to other tools.  All of the parsers reviewed can produce a
report of greater or lesser detail and complexity, a facility not offered
by \parsername{}; reporting is deferred to a subsequent application.

>>>>

@@

    \item [AWStats] \textit{AWStats is a free powerful and featureful tool
        that generates advanced web, streaming, ftp or mail server
        statistics, graphically. This log analyzer works as a CGI or from
        command line and shows you all possible information your log
        contains, in few graphical web pages. It uses a partial information
        file to be able to process large log files, often and quickly. It
        can analyze log files from all major server tools like Apache log
        files (NCSA combined/XLF/ELF log format or common/CLF log format),
        WebStar, IIS (W3C log format) and a lot of other web, proxy, wap,
        streaming servers, mail servers and some ftp servers.\/}

        AWStats will produce simple graphs for many different services, but
        that restricts it to supporting the Lowest Common Denominator.  The
        data it will extract from an MTA log file is limited to: \newline
        \hspace*{2em} time2, email, email\_r, host, host\_r, method, url,
        code and bytesd.  \newline  There does not seem to be an
        explanation of those fields in the documentation.  AWStats has no
        special purpose code to deal with the intricacies of Postfix logs,
        in fact it operates by coercing Postfix logs into
        Apache\footnote{The Apache web server is the most popular HTTP
        server in use for the past 10 years; more information is available
        at \url{http://httpd.apache.org/}} format log files, for analysis
        by its HTTP log file parser.  The converting parser is hard coded,
        hard to extend, and only deals with a small portion of the log
        lines generates by Postfix, but it does correlate log lines by
        queueid.

        \url{http://awstats.sourceforge.net/} \newline
        \url{http://awstats.sourceforge.net/awstats.mail.html} \newline
        \url{http://awstats.sourceforge.net/docs/awstats_faq.html#MAIL}
        \newline (Last checked 2007/10/13.)

@@

    \item [Logrep] \textit{Logrep is a secure multi-platform framework for
        the collection, extraction, and presentation of information from
        various log files. It features HTML reports, multi dimensional
        analysis, overview pages, SSH communication, and graphs, and
        supports over 30 popular systems including Snort, Squid, Postfix,
        Apache, Sendmail, syslog, ipchains, iptables, NT event logs,
        Firewall-1, wtmp, xferlog, Oracle listener and Pix.\/}

        Logrep extracts the following fields from Postfix logs:

            \item For mail sent and received: from address, size, and
                time and date.

            \item For mail sent: to addresses, SMTP code, and delay.

            \item For mail received: the hostname of the sender.

        Fields are correlated based on the queueid (referred to as
        sessionname within Logrep), but not by pid.  The parsing is hard
        coded, hard to extend, and error prone: empty fields are saved when
        the log line doesn't match the regex, though it appears that they
        will not overwrite existing data.  It also counts the number of log
        lines parsed and skipped.  Most notably rejections are completely
        ignored, making it unsuitable for the purposes of this project.
        Needless to say it doesn't attempt to address any of the
        complications in sections~\ref{complications} and~\ref{additional
        complications} except for correlating by queueid.

        Logrep does not come with any documentation, though some scanty
        documentation is available on its website.  It requires a web
        browser to interact with it, ensuring that automated log processing
        will be difficult.  Sadly all attempts to use Logrep failed, as it
        was unable to access the log files selected; this appears to be a
        bug rather than operator error.  As parsing failed it wasn't
        possible to review the reports Logrep can generate (available in
        HTML only), nor to examine the (undocumented) format in which it
        can save extracted data for subsequent reuse.

        Logrep extracts far less data from Postfix logs than \parsername{},
        completely ignores rejections, is effectively undocumented, doesn't
        deal with the more complicated aspects of Postfix logs, cannot be
        easily extended by the user, and at the time of writing doesn't
        work properly.

        \url{http://www.itefix.no/phpws/index.php} \newline (Last checked
        2007/10/13.)

    \item [Log Mail Analyser] Please see the previous discussion in
        section~\ref{prior art}.

@@

While there are other programs available which perform basic Postfix log
parsing (some to a greater level of detail than others), few
attempt to correlate log lines queueid (none correlate by pid) to produce
an overall record of the journey each mail traverses through Postfix.  No
other parser collects the breadth of information gathered by \parsername{},
nor makes it as easy to add new parsing rules.  Most other parsers
immediately generate a report and discard the data extracted from the logs;
those which don't discard the data retain it in a format inaccessible to
other tools.  All of the parsers reviewed can produce a report of greater
or lesser detail and complexity, a facility not offered by \parsername{};
reporting is deferred to a subsequent application.
